---
title: "Explaining the Helpers"
output: html_document
date: "2023-11-25"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
load packages
```{r, message=FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(sf)
library(here)
library(nngeo)
library(units)
library(knitr)
library(RColorBrewer)
library(leaflet)
library(dplyr)

here("code",
     "helpers.R") |>
  source()
```

load data
```{r message=FALSE, results='hide'}
parcels <- here("data",
                "test-parcels.geojson") |>
  st_read()

streets <- here("data",
                "test-streets.geojson") |>
  st_read()
```

# fun bearing
original in helpers.r
```{r, eval=FALSE}
fun_bearing = function(my_line) {
  
  if(st_is_longlat(my_line)) {
    stop("Please use a projected coordinate system")
  }
  
  coords <- st_coordinates(my_line) |>
    as_tibble() |>
    group_by(L1) |>
    summarise(height = max(X) - min(X),
              width = max(Y) - min(Y)) |>
    mutate(bearing = atan(width/height)*180/pi)
  
  coords$bearing
}
```

So, going through the functions in the helpers document, I gather that
fn_bearing finds the 'angle' of the streets
I took out the coords$bearing so I could see the whole tibble
```{r}
streetsproj <- streets |>
  st_as_sf()|>
  st_transform(26913)

streetbear  <- fun_bearing(streetsproj)

streetbear|>
  kable()|>
  head()
```
this is where bearing = atan(width/height)*180/pi and
height = max(X) - min(X) and  width = max(Y) - min(Y))
remembering that this is on a projected coordinate ref, the length and height are imagined values parallel and perpendicular to n/s, and the tan(width/height) is going to determine the angle of the actual line in relation to the line which is n/s parallel. 

# fn_min diff

I'm less sure as to what fn_min_diff is doing

```{r, eval=FALSE}
fn_min_diff <- function(list_nums, num) {
  min(abs(num - unlist(list_nums)))
}
```

}
breaking this up: 
?min()
max and min return the maximum or minimum of all the values present in their arguments, as integer if all are logical or integer, as double if all are numeric, and character otherwise.
?abs()
abs(x) computes the absolute value of x, sqrt(x) computes the (principal) square root of x...
so, it looks like it is straightforwardly selecting the smallest value of (num-list_nums), I wonder why?...

```{r}
fn_min_diff(streetbear$L1, streetbear$bearing)
```


# fn_label sides

Moving on to the BIG function...
fn_label_sides !!!!


Code in helpers.R
```{r, eval=FALSE}
fn_label_sides <- function(my_parcels,
                        my_streets,
                        close_threshold = 3,
                        parallel_threshold = 15) {
  
  edges <- my_parcels |>
    st_segments(progress = FALSE) 
```
How does it work?
```{r}
edges <- parcels |>
    st_segments(progress = FALSE)

edges|>
  head()|>
  kable()
```
?st_segments
Split lines or polygons to separate segments. lets see how this works
although this particular Kable()|> head() is ugly, it shows that the function breaks up the polygon into lines that each have their own row, but retain the OBJECTID (amongst other info) of their parent shape

Code in helpers.R
```{r, eval=FALSE}
  edges$edge_id <- seq(1:nrow(edges))
```

How does it work?
```{r}
edges1 <- edges

edges1$edge_id <- seq(1:nrow(edges))
```

so this adds a column at the end that lists all of the new segments numerically

Code in helpers.R
```{r, eval=FALSE} 
  shared <- st_equals(edges, edges, remove_self = TRUE) |>
    sapply(length)
  
  edges$unique = (shared == 0)
```

How does it work? note; I selected important columns for legibility
```{r} 
edges2 <- edges1 |>
  select(OBJECTID, PID, LAT, LON, ShapeSTArea, ShapeSTLength, result, edge_id)

  shared <- st_equals(edges2, edges2, remove_self = TRUE) |>
    sapply(length)
  
  edges2$unique = (shared == 0)
  
edges2|>
head()|>
kable()
  
```

This adds a column at the end of the tibble 'Unique' which equals TRUE or FALSE
The syntax is hard for me to understand


Code in helpers.R 
```{r, eval=FALSE} 
  street_segs <- st_segments(streets, progress = FALSE)
  
  close_street_index <- st_nn(edges, street_segs, 
                              k = 2, progress = FALSE)
  
  close_street_index_df <- matrix(unlist(close_street_index),
                                  ncol=2,
                                  byrow=TRUE) |>
    as_tibble() 
```

what does it do?* (this takes >5 minutes to run)
```{r, message= FALSE} 
  street_segs <- st_segments(streets, progress = FALSE)
  
  close_street_index <- st_nn(edges2, street_segs, 
                              k = 2, progress = FALSE)
  
  close_street_index_df <- matrix(unlist(close_street_index),
                                  ncol=2,
                                  byrow=TRUE) |>
    as_tibble() 
```

This bit cuts the street up into sections, then finds the nearest lot section to each of the respective streets
k = 2 means that it finds the 2 nearest neighbors. st_nn returns as a Large List, so matrix() turns that list into a lovely little matrix with 2 columns. 
unsure of what unlist does exactly **

```{r}
close_street_index_df |>
  head()|>
  kable()
```

Code in helpers.R
```{r, eval = FALSE}
  close_street_1 <- street_segs[close_street_index_df$V1,] 
  close_street_2 <- street_segs[close_street_index_df$V2,] 
```

what does it do?*
```{r}
  close_street_1 <- street_segs[close_street_index_df$V1,] 
  close_street_2 <- street_segs[close_street_index_df$V2,] 

close_street_1 |>
  head()|>
  kable()
```
Ok, this returns as tibbles with 67 columns.. unsure of how all these are nested in close_street_index_df
Is close_street_index not just a list of integers? This is a data structure*

Code in helpers.r
```{r, eval= FALSE}

  edges$street_bearing_1 <- fun_bearing(st_transform(close_street_1, 2812))
  edges$street_bearing_2 <- fun_bearing(st_transform(close_street_2, 2812))
  edges$edge_bearing <- fun_bearing(st_transform(edges, 2812))
  
```

what does it do?
```{r}
edges3 <- edges2

  edges3$street_bearing_1 <- fun_bearing(st_transform(close_street_1, 2812))
  edges3$street_bearing_2 <- fun_bearing(st_transform(close_street_2, 2812))
  edges3$edge_bearing <- fun_bearing(st_transform(edges3, 2812))

edges3 |>
  head()|>
  kable()
```
adding the bearings figured in fun_bearing to the df

this big one....

```{r, eval=FALSE} 

  
  edges_labeled <- edges |>
    mutate(angle_1 = abs(edge_bearing - street_bearing_1),
           angle_2 = abs(edge_bearing - street_bearing_2)) |>
    mutate(dist_to_para_st = ifelse(angle_1 < parallel_threshold,
                                    st_distance(edges,close_street_1,by_element = TRUE),
                                    st_distance(edges,close_street_2, by_element = TRUE))) |>
    group_by(PID) |>
    mutate(min_dist = min(dist_to_para_st)) |>
    mutate(diff = dist_to_para_st - min_dist) |>
    mutate(front = diff <= close_threshold &
             unique) |>
    mutate(nfronts = sum(front),
           front_bearings = list(edge_bearing[front])) |>
    ungroup() |>
    mutate(angle_from_front = mapply(fn_min_diff, front_bearings, edge_bearing)) |>
    mutate(side = case_when(front ~ "front",
                            angle_from_front < parallel_threshold ~ "rear",
                            TRUE ~ "side")) |>
    select(PID, side) |>
    group_by(PID, side) |>
    summarise(geometry = st_union(result))
}

```

```{r,} 
  edges4 <- edges3 |>
    mutate(angle_1 = abs(edge_bearing - street_bearing_1),
           angle_2 = abs(edge_bearing - street_bearing_2)) 
edges4 |>
  head()|>
  kable()
```
This is finding the difference between the angle of the 2 nearest streets and the angle of the parcel lines,
(seeing which is most parallel)

For this under, the original had parallel_threshold=15, so I just subbed 15
```{r,}
edges5 <- edges4 |>
    mutate(dist_to_para_st = ifelse(angle_1 < 15,
                                    st_distance(edges,close_street_1,by_element = TRUE),
                                    st_distance(edges,close_street_2, by_element = TRUE))) 

edges5 |>
  head()|>
  kable()
```
If the angle_1 is less than the parallel threshold, the dist_to_para_st will be the distance between close_st_1 and the street
if else, the dist_to_para_st will be the distance between the close_st_2 and the nearest street

```{r,}
edges6 <- edges5 |>
    group_by(PID)

edges6 |>
  head()|>
  kable()
```
group by PID, simple enough. I wonder where the group is 'stored' and how to refer to it **

```{r,}
edges7 <- edges6|>
    mutate(min_dist = min(dist_to_para_st)) |>
    mutate(diff = dist_to_para_st - min_dist)

edges7 |>
  head()|>
  kable()
```
min dist is the smallest dist(dist_to _para_st) within the PID group,
then the diff = dist_to_para_st - min_dist
unsure of how groups operate *

close_threshhold = 3, I subbed
```{r,}
edges8 <- edges7|>
    mutate(front = diff <= 3 &
             unique) |>
    mutate(nfronts = sum(front),
           front_bearings = list(edge_bearing[front])) 
edges8 |>
  head()|>
  kable()
```
front syntax is hard for me to understand *
n front is the number of 'front = true' in each PID group

```{r,}
edges9 <- edges8 |>
    ungroup() |>
    mutate(angle_from_front = mapply(fn_min_diff, front_bearings, edge_bearing)) 

edges9 |>
  head()|>
  kable()
```


15 sub for parallel threshold
```{r,}
edges10 <- edges9 |>
    mutate(side = case_when(front ~ "front",
                            angle_from_front < 15 ~ "rear",
                            TRUE ~ "side"))
edges10 |>
  head()|>
  kable()
```
unsure of what mapply() did, so unsure of how to interpret ***

This last one is straightforward
```{r,}
edgeslabelled <- edges10 |>
    select(PID, side) |>
    group_by(PID, side) |>
    summarise(geometry = st_union(result))

edgeslabelled |>
  head()|>
  kable()
```
  